---
title: "IQM Degradation of Deep Learning Reconstructed Prostate MRI an Evaluation of Lesion Regions."
author: "Quintin van Lohuizen"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Loading necessary libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
library(car)
library(lmerTest)
library(emmeans)
```



# Loading Data
```{r loading_data}
data_path <- "C:/Users/Quintin/Documents/phd_local/02_repos/mri-dl-recon-fov-analysis/data/final/iqms_vsharp_r1r3r6_with_ref_regions.csv"
raw_data <- read_csv2(data_path, locale = locale(decimal_mark = "."), col_types = cols(
  pat_id = col_character(),
  acceleration = col_character(),
  roi = col_character(),
  slice = col_double(),
  ssim = col_character(),
  psnr = col_character(),
  rmse = col_character(),
  hfen = col_character()
))
summary(raw_data)
str(raw_data)
```



# Preprocessing Data
```{r preprocessing}

# Convert columns to appropriate types
iqms_data <- raw_data %>%
  mutate(
    pat_id = as.factor(pat_id),
    acceleration = factor(as.numeric(acceleration), levels = c(3.0, 6.0), labels = c("3x", "6x")),
    roi = as.factor(roi),
    slice = as.numeric(slice),
    ssim = as.numeric(ssim),
    psnr = as.numeric(psnr),
    rmse = as.numeric(rmse),  # Ensure rmse is read correctly
    hfen = as.numeric(hfen)
  )

# Check for NA values and handle them if necessary
iqms_data <- iqms_data %>%
  drop_na()

# Summary to check changes
summary(iqms_data)
str(iqms_data)

# Proper rounding for numeric columns (if needed)
iqms_data <- iqms_data %>%
  mutate(
    ssim = round(ssim, 3),
    psnr = round(psnr, 2),
    rmse = round(rmse, 2),
    hfen = round(hfen, 3)
  )

# Display the first few rows of the preprocessed data
head(iqms_data)
```



# Descriptive Statistics
Calculate mean, median, standard deviation, and range for each image quality metric (SSIM, PSNR, RMSE, HFEN) across different acceleration factors and ROIs.
```{r summary_stats}
# Calculate summary statistics for each image quality metric by acceleration and ROI
summary_stats <- iqms_data %>%
  group_by(acceleration, roi) %>%
  summarise(
    mean_ssim = mean(ssim, na.rm = TRUE),
    median_ssim = median(ssim, na.rm = TRUE),
    sd_ssim = sd(ssim, na.rm = TRUE),
    range_ssim = range(ssim, na.rm = TRUE),

    mean_psnr = mean(psnr, na.rm = TRUE),
    median_psnr = median(psnr, na.rm = TRUE),
    sd_psnr = sd(psnr, na.rm = TRUE),
    range_psnr = range(psnr, na.rm = TRUE),

    mean_rmse = mean(rmse, na.rm = TRUE),
    median_rmse = median(rmse, na.rm = TRUE),
    sd_rmse = sd(rmse, na.rm = TRUE),
    range_rmse = range(rmse, na.rm = TRUE),

    mean_hfen = mean(hfen, na.rm = TRUE),
    median_hfen = median(hfen, na.rm = TRUE),
    sd_hfen = sd(hfen, na.rm = TRUE),
    range_hfen = range(hfen, na.rm = TRUE)
  ) %>%
  ungroup()

# Display the summary statistics
summary_stats
```



## Visualization
### Boxplot
Visualize the distribution of each metric across different acceleration factors and ROIs.
```{r boxplot}
# Define the full name map for ROIs
full_name_map <- c(
  'FAV' = 'Full Abdominal View (FAV, ',
  'CPV' = 'Clinical Prostate View (CPV, ',
  'TLV' = 'Targeted Lesion View (TLV, ',
  'PR' =  'Prostate Region (PR, ',
  'FR' =  'Femur Region (FR, ',
  'MR' =  'Muscle Region (MR, ',
  'SFR' = 'Sub. Fat Region (SFR, '
)
print(full_name_map)

# Calculate the number of samples for each ROI
roi_sample_sizes <- iqms_data %>%
  group_by(roi) %>%
  summarise(n = n())

# Ensure the ordering of roi_sample_sizes matches full_name_map
roi_sample_sizes <- roi_sample_sizes %>%
  arrange(factor(roi, levels = names(full_name_map)))

# Update the full name map with sample sizes
full_name_map <- setNames(
  paste(full_name_map, "n=", roi_sample_sizes$n, ")", sep = ""),
  roi_sample_sizes$roi
)

# Reorder the levels for 'roi' factor with the specified order
iqms_data$roi <- factor(iqms_data$roi, levels = c('FAV', 'CPV', 'TLV', 'PR', 'FR', 'MR', 'SFR'))

# Define the metrics and their respective y-axis limits
metrics_limits <- list(
  ssim = c(0.22, 1.0),
  hfen = c(0.1, 1.5),
  psnr = c(9, 43),
  rmse = c(0, 70)
)

# Loop through each metric and create the boxplot with the specified y-axis limits
for (metric in names(metrics_limits)) {
  limits <- metrics_limits[[metric]]
  p <- ggplot(iqms_data, aes(x = acceleration, y = .data[[metric]], fill = roi)) +
    geom_boxplot(outlier.shape = NA) +  # remove outliers from the plot
    theme_minimal() +
    theme(legend.position = "right") +
    labs(title = paste("Boxplot of", toupper(metric)), y = toupper(metric), x = "Acceleration Factor") +
    scale_fill_brewer(palette = "Set1", name = "Region of Interest", labels = full_name_map[levels(iqms_data$roi)]) +
    coord_cartesian(ylim = limits)  # adjust the y-axis limits based on the metric
  
  # Save the plot to file
  ggsave(filename = paste0("boxplot_", metric, ".png"), plot = p, width = 8, height = 6, dpi = 300, bg = "white")
  
  print(p)
}
```


# Statistical Analysis
## Data Preparation
```{r data_prep_stats}
# lets filter on a specific acceleration
iqms_data_3x <- iqms_data %>% filter(acceleration == "3x")
iqms_data_6x <- iqms_data %>% filter(acceleration == "6x")
```



## Comparative Analysis
### Wilcoxon Signed-Rank Test - Assumptions
The following histograms illustrate the distribution of SSIM values for different fields of view (FOVs). Specifically, the comparisons are between the Full Abdominal View (FAV) and Targeted Lesion View (TLV), Clinical Prostate View (CPV) and TLV, and TLV and Prostate Region (PR).

Check the density plots for each comparison to ensure that the shapes of the distributions are similar, which is necessary for the Wilcoxon Signed-Rank Test. Similar distribution shapes across groups suggest that the test's assumptions are satisfied.
```{r wilcoxon_assumptions}
# Check distribution shapes for the groups
plot_density <- function(data, metric, roi1, roi2, color1, color2) {
  filtered_data <- data %>% filter(roi %in% c(roi1, roi2))
  ggplot(filtered_data, aes_string(x = metric, fill = "roi")) +
    geom_density(alpha = 0.5) +
    labs(title = paste("Density Plot:", roi1, "vs.", roi2), x = toupper(metric), y = "Density") +
    scale_fill_manual(values = setNames(c(color1, color2), c(roi1, roi2))) +
    theme_minimal()
}
# Density plots for SSIM
plot_density(iqms_data_3x, "ssim", "FAV", "TLV", "blue", "red")
plot_density(iqms_data_3x, "ssim", "CPV", "TLV", "blue", "red")
plot_density(iqms_data_3x, "ssim", "TLV", "PR", "blue", "red")

plot_density(iqms_data_3x, "hfen", "FAV", "TLV", "blue", "red")
plot_density(iqms_data_3x, "hfen", "CPV", "TLV", "blue", "red")
plot_density(iqms_data_3x, "hfen", "TLV", "PR", "blue", "red")

print("Check density plots for similar distribution shapes for SSIM.")
```
### Mann-Whitney actually - Wilcoxon Signed-Rank Test (also known as the Mann-Whitney U Test when comparing independent samples)
Since the comparisons involve independent groups (e.g., different ROIs for potentially different slices or patients), using the unpaired version of the test (Mann-Whitney U Test) is correct.
The Wilcoxon Signed-Rank Test is used to compare paired or matched samples, while the Mann-Whitney U Test is for independent samples, which fits your dataset structure.
```{r mann_whitney}
perform_mann_whitney <- function(data, metric, roi1, roi2) {
  filtered_data <- data %>% filter(roi %in% c(roi1, roi2))
  test_result <- wilcox.test(as.formula(paste(metric, "~ roi")), data = filtered_data)
  
  print(paste("Mann-Whitney U Test for", toupper(metric), "between", roi1, "and", roi2))
  print(test_result)
  cat("------------------------------------------------\n")
  
  return(test_result)
}

# Mann-Whitney U Test for SSIM
fav_tlv_ssim_test <- perform_mann_whitney(iqms_data_3x, "ssim", "FAV", "TLV")
cpv_tlv_ssim_test <- perform_mann_whitney(iqms_data_3x, "ssim", "CPV", "TLV")
tlv_pr_ssim_test  <- perform_mann_whitney(iqms_data_3x, "ssim", "TLV", "PR")

# Mann-Whitney U Test for HFEN
fav_tlv_hfen_test <- perform_mann_whitney(iqms_data_3x, "hfen", "FAV", "TLV")
cpv_tlv_hfen_test <- perform_mann_whitney(iqms_data_3x, "hfen", "CPV", "TLV")
tlv_pr_hfen_test  <- perform_mann_whitney(iqms_data_3x, "hfen", "TLV", "PR")
```



## Correlation Analysis
### Spearman's Rank Correlation - Accelerations
```{r spearman_accs}
# Convert acceleration factor to numeric by removing "x"
iqms_data <- iqms_data %>%
  mutate(acceleration_numeric = as.numeric(gsub("x", "", acceleration)))

# Calculate Spearman's Rank Correlation for acceleration factor and image quality metrics
spearman_corr_accel <- iqms_data %>%
  select(acceleration_numeric, ssim, psnr, rmse, hfen) %>%
  cor(method = "spearman")

# Print the correlation matrix
print("Spearman's Rank Correlation Matrix including Acceleration:")
print(spearman_corr_accel)
```



### Spearman's Rank Correlation - IQMs
Assess the correlation between different image quality metrics (e.g., SSIM vs. PSNR) and between metrics and acceleration factors.
```{r spearman_iqms}
# Calculate Spearman's Rank Correlation
spearman_corr <- iqms_data_3x %>%
  select(ssim, psnr, rmse, hfen) %>%
  cor(method = "spearman")

# Print the correlation matrix
print("Spearman's Rank Correlation Matrix for 3x Acceleration:")
print(spearman_corr)

# Repeat for 6x acceleration
spearman_corr_6x <- iqms_data_6x %>%
  select(ssim, psnr, rmse, hfen) %>%
  cor(method = "spearman")

print("Spearman's Rank Correlation Matrix for 6x Acceleration:")
print(spearman_corr_6x)

```
**Summary:**
- SSIM and PSNR consistently show a strong positive correlation across both acceleration factors, indicating they generally move together.
- SSIM and HFEN have a strong negative correlation in both cases, suggesting that as the structural similarity improves, the high-frequency error norm decreases.
- PSNR and HFEN also have a strong negative correlation, meaning higher PSNR values correspond to lower HFEN values, indicating better image quality.
- RMSE shows weak and inconsistent correlations with other metrics, indicating it may be less sensitive or relevant in capturing the same aspects of image quality as SSIM, PSNR, and HFEN.


### Scatter Plots - IQMs
Visualize relationships between image quality metrics and acceleration factors.
```{r scatter_correlation}
# Function to create scatter plots
plot_scatter <- function(data, x_metric, y_metric, color_var, acc) {
  ggplot(data, aes_string(x = x_metric, y = y_metric, color = color_var)) +
    geom_point(alpha = 0.6) +
    geom_smooth(method = "lm", se = FALSE, color = "black") +
    labs(title = paste("Scatter Plot:", toupper(y_metric), "vs.", toupper(x_metric), "on", acc),
         x = toupper(x_metric),
         y = toupper(y_metric)) +
    theme_minimal() +
    scale_color_brewer(palette = "Set1")
}

# Scatter plots for 3x acceleration
# plot_scatter(iqms_data_3x, "ssim", "psnr", "roi")
# plot_scatter(iqms_data_3x, "ssim", "rmse", "roi")
plot_scatter(iqms_data_3x, "ssim", "hfen", "roi", "3x")

# Scatter plots for 6x acceleration
# plot_scatter(iqms_data_6x, "ssim", "psnr", "roi")
# plot_scatter(iqms_data_6x, "ssim", "rmse", "roi")
plot_scatter(iqms_data_6x, "ssim", "hfen", "roi", "6x")

```



## Regression Analysis
### Linear Mixed-Effects Models with CPV as Baseline
Model the relationship between SSIM and acceleration factors, accounting for random effects due to patient variability and slice index.

```{r lmem_cpv_baseline}
# Relevel the ROI factor to make CPV the baseline
iqms_data <- iqms_data %>%
  mutate(roi = relevel(roi, ref = "CPV"))

# Fit the linear mixed-effects model with CPV as the baseline
lme_model_ssim_cpv <- lmer(ssim ~ acceleration * roi + (1 | pat_id/slice), data = iqms_data)
summary(lme_model_ssim_cpv)
```
### Interpretation of Linear Mixed-Effects Model Results with CPV as Baseline
**Fixed Effects**:  
- **Intercept (0.8766)**: The baseline SSIM value for CPV with 3x acceleration.  
- **Acceleration (6x: -0.0212)**: The change in SSIM when the acceleration factor is 6x compared to 3x for CPV. The negative estimate (-0.0212) indicates a significant reduction in SSIM for 6x acceleration.  
- **ROI Effects**:  
  - **FAV (0.0230)**: SSIM for FAV is higher than CPV by 0.0230.  
  - **TLV (-0.2168)**: SSIM for TLV is lower than CPV by 0.2168.  
  - **PR (-0.2585)**: SSIM for PR is lower than CPV by 0.2585.  
  - **FR (-0.2142)**: SSIM for FR is lower than CPV by 0.2142.  
  - **MR (-0.1865)**: SSIM for MR is lower than CPV by 0.1865.  
  - **SFR (-0.0923)**: SSIM for SFR is lower than CPV by 0.0923.  
- **Interaction Effects**:
  - **FAV (0.0203)**: SSIM for FAV with 6x acceleration is higher by 0.0203 compared to CPV.  
  - **TLV (-0.0365)**: SSIM for TLV with 6x acceleration is lower by 0.0365 compared to CPV.  
  - **PR (-0.0619)**: SSIM for PR with 6x acceleration is lower by 0.0619 compared to CPV.  
  - **FR (-0.1116)**: SSIM for FR with 6x acceleration is lower by 0.1116 compared to CPV.  
  - **MR (-0.0445)**: SSIM for MR with 6x acceleration is lower by 0.0445 compared to CPV.  
  - **SFR (-0.0397)**: SSIM for SFR with 6x acceleration is lower by 0.0397 compared to CPV.  

**Random Effects**:
- **Patient ID (Variance: 0.0036, Std.Dev: 0.0597)**: Indicates variability in SSIM due to different patients.  
- **Slice within Patient (Variance: 0.0007, Std.Dev: 0.0261)**: Indicates variability in SSIM due to different slices within the same patient.  
- **Residual (Variance: 0.0071, Std.Dev: 0.0844)**: Residual variability not explained by the model.  

### Critical Analysis:  

**Baseline Adjustment**:  
- Setting CPV as the baseline makes the results more relevant for practical applications, as CPV is the standard for reconstruction evaluation.  
- The intercept (0.8766) indicates the baseline SSIM for CPV at 3x acceleration is robust and high.  

**Acceleration Effect**:
- The negative coefficient for acceleration (6x: -0.0212) suggests that increasing acceleration to 6x reduces SSIM for CPV significantly, highlighting a potential trade-off between acceleration and image quality.  

**ROI Effects**:
- TLV, PR, FR, MR, and SFR all show significantly lower SSIM compared to CPV, with PR having the largest negative effect (-0.2585).  
- FAV shows a slight positive effect on SSIM compared to CPV (0.0230), indicating better performance.  

**Interaction Effects**:
- The interaction terms reveal that the effect of increasing acceleration to 6x is different across ROIs.  
- For example, TLV with 6x acceleration has a further reduction in SSIM (-0.0365) compared to CPV.  
- PR, FR, MR, and SFR also show significant negative interactions with 6x acceleration, with FR having the largest negative interaction (-0.1116).  

**Clinical Relevance**:
- The significant reduction in SSIM for TLV and PR with higher acceleration factors is critical, as these regions are essential for accurate lesion detection and diagnosis.  
- The findings suggest that while acceleration improves efficiency, it may compromise image quality in critical regions, necessitating careful balancing in clinical settings.  

**Model Fit and Residuals**:
- The model fit appears reasonable given the random effects, but the residual variance suggests there are other factors influencing SSIM not captured by the model.  
- Further model diagnostics (e.g., residual plots) could help identify any patterns or violations of assumptions.  

**Conclusion**:
- The results indicate a significant impact of acceleration on SSIM, especially in critical ROIs like TLV and PR.


### Model Diagnostics: Residual Plots
To identify any patterns or violations of assumptions, we can plot the residuals of the model. Residual plots can help detect non-linearity, heteroscedasticity, and outliers.
```{r model_diagnostics}
# Extract residuals and fitted values from the model
residuals <- resid(lme_model_ssim_cpv)
fitted_values <- fitted(lme_model_ssim_cpv)

# Create a residuals vs. fitted values plot
ggplot(data.frame(fitted_values, residuals), aes(x = fitted_values, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs. Fitted Values",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

# Create a Q-Q plot of residuals to check normality
qqnorm(residuals)
qqline(residuals, col = "red", linetype = "dashed")

# Create a histogram of residuals to check for normal distribution
ggplot(data.frame(residuals), aes(x = residuals)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "blue", alpha = 0.7) +
  labs(title = "Histogram of Residuals",
       x = "Residuals",
       y = "Frequency") +
  theme_minimal()

# Check for heteroscedasticity using a scale-location plot
scale_location_plot <- data.frame(
  sqrt_abs_residuals = sqrt(abs(residuals)),
  fitted_values = fitted_values
)

ggplot(scale_location_plot, aes(x = fitted_values, y = sqrt_abs_residuals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", col = "red") +
  labs(title = "Scale-Location Plot",
       x = "Fitted Values",
       y = "Sqrt(|Residuals|)") +
  theme_minimal()
```


### Model Diagnostics: Residual Plots - Lets try transforming the dependent variable SSIM
```{r model_diagnostics_dep_trans}
# Apply logit transformation to SSIM values
iqms_data <- iqms_data %>%
  mutate(ssim_logit = log(ssim / (1 - ssim)))

# Fit the linear mixed-effects model with transformed SSIM
lme_model_ssim_logit <- lmer(ssim_logit ~ acceleration * roi + (1 | pat_id/slice), data = iqms_data)
summary(lme_model_ssim_logit)

# Extract residuals and fitted values from the transformed model
residuals_logit <- resid(lme_model_ssim_logit)
fitted_values_logit <- fitted(lme_model_ssim_logit)

# Residuals vs. Fitted Values Plot for the transformed model
ggplot(data.frame(fitted_values_logit, residuals_logit), aes(x = fitted_values_logit, y = residuals_logit)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs. Fitted Values (Logit Transformed SSIM)",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

# Q-Q Plot for the transformed model
qqnorm(residuals_logit)
qqline(residuals_logit, col = "red", linetype = "dashed")

# Histogram of Residuals for the transformed model
ggplot(data.frame(residuals_logit), aes(x = residuals_logit)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "blue", alpha = 0.7) +
  labs(title = "Histogram of Residuals (Logit Transformed SSIM)",
       x = "Residuals",
       y = "Frequency") +
  theme_minimal()

# Scale-Location Plot for the transformed model
scale_location_plot_logit <- data.frame(
  sqrt_abs_residuals_logit = sqrt(abs(residuals_logit)),
  fitted_values_logit = fitted_values_logit
)
ggplot(scale_location_plot_logit, aes(x = fitted_values_logit, y = sqrt_abs_residuals_logit)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", col = "red") +
  labs(title = "Scale-Location Plot (Logit Transformed SSIM)",
       x = "Fitted Values",
       y = "Sqrt(|Residuals|)") +
  theme_minimal()
```


### Slope P-values for Interaction Effects
To assess whether the slopes for 3x to 6x acceleration on SSIM are significantly different between CPV and TLV, we use a linear mixed-effects model with log-transformed SSIM as the dependent variable. The car::Anova function allows us to perform Type III Wald chi-square tests to evaluate the significance of the interaction effects between acceleration factors and ROIs.
```{r slope_pvalue_interactions_effects}

# Fit the model with log-transformed dependent variable if not already done
lme_model_ssim_cpv_log <- lmer(log(ssim) ~ acceleration * roi + (1 | pat_id/slice), data = iqms_data, REML = TRUE, control = lmerControl(optimizer = "bobyqa"))

# Use car::Anova to perform Type III Wald chi-square tests
anova_results <- Anova(lme_model_ssim_cpv_log, type = 3)
print(anova_results)
```
This doesnt seem so usefull for seeing differences between different ROIs.


# Alternative Approach: Explicit Interaction Term Creation
To make the coding explicit and ensure correct hypothesis testing, we'll create an explicit interaction term between acceleration and ROI. This approach can help avoid naming issues and facilitate direct comparison.
```{r explicit_interaction_term_creation}
# Create an explicit interaction term
iqms_data <- iqms_data %>%
  mutate(acc_roi = interaction(acceleration, roi))

# Fit the linear mixed-effects model with log-transformed SSIM and the explicit interaction term
lme_model_ssim_interaction <- lmer(log(ssim) ~ acc_roi + (1 | pat_id/slice), data = iqms_data)
summary(lme_model_ssim_interaction)

# Extract coefficients for interaction terms
coefficients_interaction <- summary(lme_model_ssim_interaction)$coefficients
coefficients_interaction

# Perform Wald test to compare slopes for CPV and TLV
# Here we test if the difference in slopes for 3x to 6x between CPV and TLV is significant
linearHypothesis(lme_model_ssim_interaction, c("acc_roi6x.TLV = acc_roi6x.CPV"))
```
**Effect of Acceleration on CPV:**
- The coefficient for acc_roi6x.CPV (-0.0246) shows a significant reduction in log(SSIM) when acceleration increases from 3x to 6x for CPV.
**Effect of Acceleration on TLV:**  
- The coefficient for acc_roi6x.TLV (-0.3829) indicates a more substantial reduction in log(SSIM) when acceleration increases from 3x to 6x for TLV, compared to CPV.  
**Difference in Slopes:**
- The Wald test confirms that the difference in slopes between CPV and TLV is highly significant, indicating that acceleration has a more detrimental effect on SSIM for TLV than for CPV.  




## Post-Hoc Analysis
### Bonferroni Correction
Adjust for multiple comparisons when performing pairwise tests to control for type I error rate.
```{r bonferroni}
# text
```
### Tukey's Honest Significant Difference Test
Use for post-hoc pairwise comparisons following ANOVA.
```{r tukey}
# text
```


---------------------------------------------------------------------------------------------------------------------------------------------------


# Empty block 2 
```{r emtpy_block2}
# text
```

# Empty block 3
```{r emtpy_block3}
# text
```